---
title: "RDA & PLS"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message="hide", warning = "hide")
```

```{r, results="hide"}
suppressMessages({
    library(mixOmics)
    library(microbenchmark)
    library(parallel)
    library(elasticnet)
    library(Rfast)
    library(parallel)
})
```


```{r}
sourceCpp("../src/utils/test.cpp")
```

```{r, results="hide"}
suppressMessages({
    file.sources = list.files(path = "../src/utils", pattern = "*.r", ignore.case = T, full.names = TRUE)
    sapply(file.sources,source)
})

```


# breast cancer data

```{r}
data("breast.TCGA")
X <- breast.TCGA$data.train$mrna
Y <- breast.TCGA$data.train$protein
```

## run time

RDA using Rcpp

```{r}
#run spls from mixOmics
res_spls <- spls(X,Y,
                 keepX = c(25, 25),
                 keepY = c(dim(Y)[2],dim(Y)[2]))

#run sRDA / sCCA from sRDA pacakge
res_sRDA <- sRDAccp(X, Y,
                    nonzero = c(25),
                    multiple_LV = T, nr_LVs = 2,
                    penalization = "ust")
```


```{r, results="hide"}
set.seed(100)
spls_bc <- expression(spls(X = X,Y = Y, keepX = c(25, 25), 
                            keepY = c(dim(Y)[2],dim(Y)[2]),
                            ncomp = 2))
rda_bc <- expression(sRDAccp(predictor = X, predicted = Y,
                              nonzero = c(25),
                              multiple_LV = T, nr_LVs = 2,
                              penalization = "ust"))
mb_out <- microbenchmark(eval(spls_bc),
               eval(rda_bc), 
               times = 100, unit = "s")
```

```{r}
mb_out
```


On the small scale data `sRDA` outperforms only slightly `spls`.
on bigger data there is a significant performance gain:
on this example sRDA is about 4 times faster than `spls`:
## bigger dataset

```{r}
set.seed(100)
A <- matrix(rnorm(1000000), 100, 10000)
B <- matrix(rnorm(10000), 100, 100)
dim(A); dim(B)
# [1]   100 10000
# [1] 100 100
```

```{r, results="hide"}

spls_bigdata <- expression(spls(X = A,Y = B, 
                            keepX = c(25, 25), 
                            keepY = c(dim(B)[2],dim(B)[2]),
                            ncomp = 2))
rda_bigdata <- expression(sRDAccp(predictor = A, predicted = B,
                              nonzero = c(25),
                              multiple_LV = T, nr_LVs = 2,
                              penalization = "ust"))

mb_big <- microbenchmark(eval(spls_bigdata),
               eval(rda_bigdata), 
               times = 10, unit = "s")
```

```{r}
mb_big
```


**NOTE:** Is convergence criteria comaparable?

# plots

after obtaining the results on the breast cancer data from mixOmics,
we can plot with the results of sRDA with mixOmics' plots too

```{r}
class(res_spls)

# after obtaining results, put sRDA outputs in mixOmics' "mixo_spls" class 
res_sRDA <- reshape_sRDA_output_to_mixOmics(mix_omics_output = res_spls,
                                            old_rda_output = res_sRDA)
```

## sRDA

```{r, fig.asp=1}
plotIndiv(res_spls)     ## sample plot     
plotVar(res_spls)       ## variable plot
plotLoadings(res_spls, comp = 1, size.name = rel(0.5))
cim(res_spls, comp = 1)
```

## sPLS

```{r, fig.asp=1}
plotIndiv(res_sRDA)
plotVar(res_sRDA)
plotLoadings(res_sRDA, comp = 1, size.name = rel(0.5))
cim(res_sRDA, comp = 1)
```

# explained variance

we can look at explained variances, they are about the same

## sPLS
```{r}
res_spls$explained_variance
```

## sRDA
```{r}
res_sRDA$explained_variance
```

# correlations of latent variates / scores

these are very similar too in both methods

## sPLS
```{r}
cor(res_spls$variates[["X"]], res_spls$variates[["Y"]])
```

## sRDA
```{r}
cor(res_sRDA$variates[["X"]], res_sRDA$variates[["Y"]])
```


# covariance

covariance is higher in sPLS and it is standardized in RDA

## sPLS
```{r}
cov(res_spls$variates[["X"]], res_spls$variates[["Y"]])
```

## sRDA
```{r}
cov(res_sRDA$variates[["X"]], res_sRDA$variates[["Y"]])
```

# sum squared correlation between latent variates and outcome

these will be the loadings at RDA and they are standardized for sPLS (equal to 1).


## sPLS

standardized and equal to one.

```{r}
sum(res_spls$loadings$Y[,1]^2)
```


## sRDA
```{r}
res_sRDA$loadings$Y[,1][1:4]
sum(res_sRDA$loadings$Y[,1]^2)
```

# CCA

```{r}
res_sCCA <- sCCA(X, Y,
                    nonzero = c(25),
                    multiple_LV = T, nr_LVs = 2,
                    penalization = "ust")

res_sCCA <- reshape_sRDA_output_to_mixOmics(mix_omics_output = res_spls,
                                            old_rda_output = res_sCCA)
```

## plots

```{r, fig.asp=1}
plotIndiv(res_sCCA)
plotVar(res_sCCA)
plotLoadings(res_sCCA, comp = 1, size.name = rel(0.5))
cim(res_sCCA, comp = 1)
```

## covariance b/w variates

```{r}
cov(res_sCCA$variates[["X"]], res_sCCA$variates[["Y"]])
```

